HGIS 2021-05-28 Carbonates

### 2021-06-14 doc/hgis_data_reduction.Rmd
Update methods.

### Read raw file

@@

### Apply 13C correction to each run of a target

@@

### Flag outliers by interquartile distance

Outliers are defined as points 1.5 * interquartile range above the 3rd quartile or below the 1st quartile, and are dropped from the analysis.

### Determine per-target means, internal and external errors

@@

### Normalize targets using mean of standards and propagate errors

@@

$$ R_{norm} = \frac{R_{sample}}{R_{std}}*1.0398$$

Normalization errors are propagated using this equation:

$$ \sigma_{norm} = \sqrt{\frac{\sigma_{sample}^2}{R_{sample}^2} + \frac{\sigma_{std}^2}{R_{std}^2}} $$

### Apply a large blank correction using the mean of blanks and propagate errors.

>>>>

## Read raw file

@@

## Apply 13C correction to each run of a target

@@

## Flag outliers by interquartile distance

Outliers are defined as points 1.5 * interquartile range above the 3rd quartile or below the 1st quartile. 25% of values are expected to lie below the first quartile and 25% above the 3rd quartile. Interquartile range (IQR) is the difference between the 3rd and 1st quartile. Outliers are not currently dropped from the analysis.

## Determine per-target means, internal and external errors

@@

## Normalize targets using mean of standards and propagate errors

@@

$$ R_{norm} = \frac{R_{sample}\cdot R_{exp}}{R_{std}}$$

If the expected ratio ($R_{exp}$) of the standard is taken as a measured value with an error, rather than a constant or assigned ratio, normalization errors are propagated using this equation:

$$ \sigma_{norm} = \sqrt{\frac{\sigma_{sample}^2}{R_{sample}^2} + \frac{\sigma_{std}^2}{R_{std}^2} +
\frac{\sigma_{exp}^2}{R_{exp}^2}} $$

## Apply a large blank correction using the mean of blanks and propagate errors.

### 2021-06-14 inst/rmd/HGIS_pendleton_2021-05-28.Rmd
Update methods.

```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE)
```

@@

- Using standards in positions `r standards` for normalization.
- Using blanks in positions `r blanks` for blank correction.
- Outliers are defined as points 1.5 * interquartile range above the 3rd quartile or below the 1st quartile, and are dropped from the analysis.

@@

1. Read raw file
1. Apply 13C correction to each run of a target
1. Flag outliers by interquartile distance
1. Determine per-target means, internal and external errors
1. Normalize targets using mean of standards and propagate errors
1. Apply a large blank correction using the mean of blanks and propagate errors.

@@

```{r}
df <- get_hgis_data(file, date, standards)
write_csv(df, here("data_analysed/HGIS_2021-05-28_raw.csv"))
write.xlsx(df, here("data_analysed/HGIS_2021-05-28_raw.xlsx"))
```

@@

```{r}
df %>% 
  filter(!(Pos %in% 1:4)) %>% 
  plot_hgis_time(normFm, ce*normFm) +
  ggtitle("Sample ratio vs. time")
```

```{r}
df %>% 
  filter(!(Pos %in% 1:4)) %>% 
  plot_hgis_time(he12C) +
  ggtitle("Sample current vs time")
```

```{r}
df_sum <- sum_hgis_targets(df) %>% 
  norm_hgis() %>% 
  blank_cor_hgis(blanks = c(6, 9, 11))

@@

```{r}
df_sum %>% 
  filter(as.numeric(Pos) > 4) %>% 
  mutate(sig_normFm = max_err * normFm) %>% 
  select(Pos, Sample.Name, he12C, n_runs, fm_corr, sig_fm_corr) %>% 
  arrange(Sample.Name) %>% 
  knitr::kable()
  # gt::gt()
```

```{r}
df_sum %>% 
  filter(as.numeric(Pos) > 4) %>% 
plot_hgis_summary()
```

@@

```{r}
df_sum %>% 
  filter(as.numeric(Pos) > 4) %>% 
  compare_replicates() %>% 
  knitr::kable()
```

@@

```{r}
cons <- df_sum %>% 
  filter(as.numeric(Pos) > 4,
         !is.na(fm_consensus)) %>% 
  select(Sample.Name, fm_consensus, fm_corr, sig_fm_corr) %>% 
  mutate(Fm_diff = fm_corr - fm_consensus,
         sigma = amstools::sigma(fm_corr, fm_consensus, sig_fm_corr)) %>% 
  arrange(Sample.Name)

>>>>

```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE, echo = FALSE)
```

@@

- Using standards in positions `r standards` for normalization.
- Using blanks in positions `r blanks` for blank correction.
- Outliers are defined as points 1.5 * interquartile range above the 3rd quartile or below the 1st quartile.
- Outliers are NOT dropped from the analysis.

@@

1. Read raw file
1. Apply 13C correction to each run of a target
1. Flag outliers by interquartile distance
1. Determine per-target means, internal and external errors
1. Normalize targets using mean of standards and propagate errors
1. Apply a large blank correction using the mean of blanks and propagate errors

@@

```{r}
df <- process_hgis_results(file, date, standards)
write_csv(df, here("data_analysed/HGIS_2021-05-28_raw.csv"))
write.xlsx(df, here("data_analysed/HGIS_2021-05-28_raw.xlsx"))
```

@@

```{r}
df %>% 
  filter(!(pos %in% 1:4)) %>% 
  plot_hgis_time(norm_ratio, sig_norm_ratio, outlier = !.$ok_calc) +
  labs(title = "Sample ratio vs. time",
       y = "Fraction modern")
```

```{r}
df %>% 
  filter(!(pos %in% 1:4)) %>% 
  plot_hgis_time(he12C * 1E6, outlier = !.$ok_calc) +
  labs(title = "Sample current vs time",
       x = "Time (s)",
       y = "Current (uA)")
```

```{r}
df_sum <- sum_hgis_targets(df, remove_outliers = FALSE) %>% 
  norm_hgis() %>% 
  blank_cor_hgis(blanks = c(6, 9, 11))

@@

```{r}
df_sum %>% 
  filter(as.numeric(pos) > 4) %>% 
  mutate(he12C = he12C * 1E6) %>% 
  select(Pos = pos, Name = sample_name, `Current (uA)` = he12C, `N runs` = n_runs, Fm = fm_corr, `Fm error` = sig_fm_corr) %>% 
  arrange(Name) %>% 
  knitr::kable()
  # gt::gt()
```

```{r}
df_sum %>% 
  filter(as.numeric(pos) > 4) %>% 
plot_hgis_summary()
```

@@

```{r}
df_sum %>% 
  filter(as.numeric(pos) > 4) %>% 
  compare_replicates() %>% 
  select(Name, `Mean current (uA)` = he12C_mean, 
         `SD of current` = he12C_sd, `Mean Fm` = fm_corr_mean, 
         `SD of Fm` = fm_corr_sd, `Mean sample error` = sig_fm_corr_mean) %>% 
  knitr::kable()
```

@@

```{r}
cons <- df_sum %>% 
  filter(as.numeric(pos) > 4,
         !is.na(fm_consensus)) %>% 
  select(sample_name, fm_consensus, fm_corr, sig_fm_corr) %>% 
  mutate(Fm_diff = fm_corr - fm_consensus,
         sigma = amstools::sigma(fm_corr, fm_consensus, sig_fm_corr)) %>% 
  arrange(sample_name) %>% 
  select(Name = sample_name, 
         #`Mean current (uA)` = he12C_mean, 
         #`SD of current` = he12C_sd, 
         `Consensus Fm` = fm_consensus,
         Fm = fm_corr, 
         `Fm error` = sig_fm_corr,
         `Difference` = Fm_diff,
         Sigma = sigma)
